---
title: "SVM"
author: "Mathieu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# clean variables
```{r}
rm(list = ls())
```


```{r}
library(tidyverse)
# color palettes
library(hrbrthemes)
library(viridis)
# train-test split
library(rsample)
# tables
library(formattable)
library(huxtable)
library(gt)
# skewness,...
library(moments)
library(patchwork)
# correlation
library(corrplot)
# modeling
library(caret)
# SVM
library(kernlab)
library(e1071)
library(pdp)
library(vip)
# PCA
library(MASS)

library(dplyr)
library(magrittr)
```


```{r}
parkinson_data <- read.csv("dataset/parkinsons.csv", sep = ",", dec = ".", header = TRUE)

parkinson_data <- parkinson_data %>%
  rename(
    mdvp.fo = MDVP.Fo.Hz.,
    mdvp.fhi = MDVP.Fhi.Hz.,
    mdvp.flo = MDVP.Flo.Hz.,
    mdvp.jitter_perc = MDVP.Jitter...,
    mdvp.jitter_abs = MDVP.Jitter.Abs.,
    mdvp.rap = MDVP.RAP,
    mdvp.ppq = MDVP.PPQ,
    jitter.ddp = Jitter.DDP,
    mdvp.shimmer = MDVP.Shimmer,
    mdvp.shimmer_db = MDVP.Shimmer.dB.,
    mdvp.apq = MDVP.APQ,
    shimmer.apq3 = Shimmer.APQ3,
    shimmer.apq5 = Shimmer.APQ5,
    shimmer.dda = Shimmer.DDA,
    nhr = NHR,
    hnr = HNR,
    rpde = RPDE,
    dfa = DFA,
    d2 = D2,
    ppe = PPE
  ) %>%
  mutate(status = factor(ifelse(status == "1", "parkinson", "healthy")))

head(parkinson_data)
```

## SVM 

#Let's split the dataset

```{r}
set.seed(123)
split <- initial_split(parkinson_data, prop = 0.7)
df_train <- training(split)
df_trainU<-upSample (df_train, df_train$status)
df_test <- testing(split)
df_testU<-upSample(df_test,df_test$status)

table(parkinson_data$status)

table(df_train$status)
table(df_trainU$status)

table(df_test$status)
table(df_testU$status)

##### note that an additional column (25) is created with the upSample function
##### to be removed when using upSample data sets: col 1 name + col 25 (col18 = status)
```


#Linear SVM with SVM function


# Optimization of cost parameter

```{r}
cost2 <- seq(0.1,15,length=30)

linear_kernel2tuneU <- tune.svm(status ~ ., data = df_trainU[,-c(1,25)], scale=TRUE, kernel='linear',  type='C-classification', cross=5, cost=cost2)                                   
                                    
summary(linear_kernel2tuneU)
```

# Use of optimized parameter

```{r}
linear_kernel2U <- svm(status~., data=df_trainU[,-c(1,25)], scale=TRUE, kernel='linear', type='C-classification', cost=linear_kernel2tuneU$best.parameters$cost)
                                    
linear_kernel2U
```


# predictions on upsample train_sample

```{r}
TrainResultU<-predict(linear_kernel2U, data=df_trainU[,-c(1,18,25)], scale=TRUE)

confus.matrixtrainU=table(real=df_trainU$status, predict=TrainResultU)

print(confus.matrixtrainU)

accuracytrainU = sum(diag(confus.matrixtrainU))/sum(confus.matrixtrainU)

cat("accuracy=",accuracytrainU)
```


# predictions on upsample test_sample

```{r}
TestResultU<-predict(linear_kernel2U, df_testU[,-c(1,18,25)], scale=TRUE)

confus.matrixtestU=table(real=df_testU$status, predict=TestResultU)

print (confus.matrixtestU)

accuracytestU =sum(diag(confus.matrixtestU))/sum(confus.matrixtestU)

cat("accuracy=",accuracytestU)
```


#objective function

```{r}
linear_kernel2U$index #index of sv
linear_kernel2U$SV #SV
linear_kernel2U$coefs 	#coeff times training labels

w<- t(linear_kernel2U$coefs) %*% linear_kernel2U$SV
b<- -linear_kernel2U$rho


sum(linear_kernel2U$coefs)  #should be 0 cf constraint

yi<- ifelse(df_trainU$status=='parkinson', 1, -1)

ObjF1term <- sum(linear_kernel2U$coefs*yi[linear_kernel2U$index])

ObjF2term <- w %*% t(w)

ObjF <- ObjF1term - 0.5 * ObjF2term
 
ObjF
```


### Kernel trick

#Optimizationof cost and gamma parameters

```{r}
kernel_svm2U<-tune.svm(status~., data=df_trainU[,-c(1,25)], scale=TRUE, kernel='radial',  type='C-classification', cross=5, cost= c(0.5, 1, 1.5,2), gamma=c(0.1,0.2,0.3,0.4,0.5,0.7,0.9,1,1.2))                                  
                                    
summary(kernel_svm2U)

kernel_svm2U$best.parameters
```

#Mathieu’s results for Cost and Gamma

```{r}
kernel_svmTunedU <- svm(status~., data=df_trainU[,-c(1,25)], scale=TRUE, kernel='radial',  type='C-classification', cost=0.455, gamma=0.357)
```

# predictions on upsample train_sample

```{r}
predRadialTrainU <- predict(kernel_svmTunedU, newdata=df_trainU[,-c(1,18,25)])

confus.matrixtrainU=table(real=df_trainU$status, predict=predRadialTrainU)

print(confus.matrixtrainU)

accuracytrainU =sum(diag(confus.matrixtrainU))/sum(confus.matrixtrainU)

cat("accuracy=",accuracytrainU)
```

# predictions on upsample test_sample

```{r}
predRadialTestU <- predict(kernel_svmTunedU, newdata=df_testU[,-c(1,18,25)])

confus.matrixtestU=table(real=df_testU$status, predict=predRadialTestU)

print(confus.matrixtestU)

accuracytestU =sum(diag(confus.matrixtestU))/sum(confus.matrixtestU)

cat("accuracy=",accuracytestU)
```

# Objective function

```{r}
kernel_svmTunedU $index #index of sv
kernel_svmTunedU $SV #SV
kernel_svmTunedU $coefs 	#coeff times training labels

n=nrow(df_trainU)


rbf_kernel <- function(x1, x2, gamma) {

  return(exp(-gamma*sum(x1-x2)^2))

}

Xx<-cbind(df_trainU[,-c(1,18,25)],1)

K <- matrix(0, n, n)

for (i in 1:n) {

  for (j in 1:n) {

    K[i,j] <- rbf_kernel(Xx[i,], Xx[j,], 0.4)

  }

}




w<- t(kernel_svmTunedU$coefs) %*% K[kernel_svmTunedU$index ,]
b<- - kernel_svmTunedU$rho



yi<- ifelse(df_trainU$status=='parkinson', 1, -1)

ObjF1term <- sum(kernel_svmTunedU$coefs*yi[kernel_svmTunedU$index])

ObjF2term <- w %*% t(w)

ObjF <- ObjF1term - 0.5 * ObjF2term

ObjF
```

# Perform PCA on data


#PCA on test sample

```{r}
x_test<-(df_test[,-c(1,18)])

x_test=as.matrix(x_test)

y1<-df_test[,18]

y2<-ifelse(y1=='parkinson',1, -1)

y_test<-as.factor(y2)

pca1 <- prcomp(x_test, center = TRUE, scale= TRUE)

dim(pca1$x)

pca_df1 <- as.data.frame(pca1$x[,1:2])

pca_df=data.frame(pca_df1, y_test)

colnames(pca_df)[3]  <- "status"

dim(pca_df)

y_pca=pca_df[,3]

X_pca=pca_df[,c(1,2)]


svm_modelPCA <- svm(y_pca ~ ., data = X_pca, kernel = "radial", cost = C_pca, gamma = gamma_pca, tol = tol_pca )

svm_modelPCA

Y_PCA<-data.frame(y_pca, X_pca)


plot (svm_modelPCA, Y_PCA, svSymbol=5, symbolPalette = rainbow(4))
```

#PCA on upsample train data

```{r}
x_train<-(df_trainU[,-c(1,18,25)])

x_train=as.matrix(x_train)

y1<-df_trainU[,18]

y2<-ifelse(y1=='parkinson',1, -1)

y_train<-as.factor(y2)

pca2 <- prcomp(x_train, center = TRUE, scale= TRUE)


x_test<-(df_testU[,-c(1,18,25)])

x_test=as.matrix(x_test)

y1<-df_testU[,18]

y2<-ifelse(y1=='parkinson',1, -1)

y_test<-as.factor(y2)

pca1 <- prcomp(x_test, center = TRUE, scale= TRUE)

dim(pca1$x)

pca_df1 <- as.data.frame(pca1$x[,1:2])

pca_df=data.frame(pca_df1, y_test)

colnames(pca_df)[3]  <- "status"

dim(pca_df)

y_pca=pca_df[,3]

X_pca=pca_df[,c(1,2)]

X_pca=scale(X_pca)

#X_pca=as.matrix(X_pca)


svm_modelPCA <- svm(y_pca ~ ., data = X_pca, kernel = "radial", cost = C_pca, gamma = gamma_pca, tol = tol_pca )

svm_modelPCA

Y_PCA<-data.frame(y_pca, X_pca)


plot (svm_modelPCA, Y_PCA, svSymbol=5, symbolPalette = rainbow(4))
```